\thetitle

### 2021-05-07 paper/figures/productionsystem.tex
update

  \path (CustomerTop) edge[densely dashed] (10.5,2.8);
  \path (10.5,2.8) edge[densely dashed,midway,align=center] node[yshift=0.0cm] {\scriptsize Customer places
    an order which initiates the process\\ {\tiny Interarrival-time distributions:
      $\Exp(135)$ or $\Unif(95,175)$ or $\Exp(118)$ or $\Unif(78,158)$ % or $\Exp(105)$ or $\Unif(65, 145)$
    }
  } (-4.5, 2.8);
  \path (-4.5,2.8) edge[densely dashed] (-4.5,0);
  \path (-4.5,0) edge[->,densely dashed] (-4.0,0);

>>>>

  \path (CustomerTop) edge[densely dashed] (10.5,2.8);
  \path (10.5,2.8) edge[densely dashed,midway,align=center] node[yshift=0.0cm] {\scriptsize Customer places
    an order which initiates the process\\ {\tiny Interarrival-time distributions:
      $\Exp(135)$ or $\Unif(95,175)$ or $\Exp(118)$ or $\Unif(78,158)$ or $\Exp(105)$ or $\Unif(65, 145)$
    }
  } (-4.5, 2.8);
  \path (-4.5,2.8) edge[densely dashed] (-4.5,0);
  \path (-4.5,0) edge[->,densely dashed] (-4.0,0);

### 2021-05-07 paper/paper.tex
update

\paragraph{Demand.} The incoming orders are placed in the order pool with a due date slack of
\dds{} periods, that is each order arriving within period $t$ will be due at the end of period
$t+\dds{}$. The due date slack of \(\dds{}\) was chosen to (i) ensure sufficient time between the
first occurrence of orders in the order pool and the latest possible release of orders and (ii) to
have a clear cause and effect relationship between the order release decision and the cost
performance in the analysis.
%
To simulate moderate, medium and high variability of the demand process the interarrival time
between consecutive incoming orders is either drawn from an exponential ($\Exp$) or an uniform
($\Unif$) distribution, which are adjusted to yield the desired bottleneck utilisation level of
either $70\%$ with $\Exp(135)$ and $\Unif(95,175)$, $80\%$ with $\Exp(118)$ and $\Unif(78,158)$, or
$90\%$ with $\Exp(105)$ and $\Unif(65,145)$. These values were chosen as they comprise the
non-linear relationship between flow times and high utilisation levels.

@@

\paragraph*{Neural Network Setup.}
%
Due to the vast possibilities the neural network architecture was manually optimised, starting with
less stochastic systems and aiming for a coffin shape that can be used in all setups. During this
process we constantly adapted and reevaluated the best network on various production system
configurations. The resulting architecture depends on the number of features \(\anninp\) and the
number of output nodes \(\annout\), which itself depends on the number of actions
\(\acts = \size{\States}\), the independent agent configuration, and the number of functions that
are approximated. For the later, in case of \ARA{} this is \(2\), as we represent both functions
\(\X^\pol_{\gamma_{0}}\) and \(\X^\pol_{\gamma_{1}}\) with the same neural network, while deep
Q-Learning only approximates the function \(Q_{\gamma}^\pol\).
%
The resulting feedforward deep ANN architecture is given as %
\(\anninp %
\rightarrow 1.75 \anninp % \rightarrow \relu
\rightarrow 1.5 \anninp % \rightarrow \relu
% \rightarrow 2 \annout % \rightarrow \relu
\rightarrow \annout %
% \rightarrow   \leakyTanh%
\), where each arrow symbolises a fully-connected layer with a leaky rectified linear unit
(\(\relu\)) as activation function, which leaks values smaller than \(0\) with rate \(0.02\).
%
Note
that we explicitly do not use an output activation function. The idea is that we do not want to
restrict the approximations to values inside this range. This is especially important for
Q-Learning, as the range of the state values are difficult to forecast.
%
The actual values \(\anninp\) and \(\annout\) depend on the used algorithm. In our case case using 6
independent agents we gain \(\anninp=324\) and \(\annout=36\). For the Q-Learning algorithm, which
only approximates one function, the output nodes \(\annout\) are halved.
%
% SpecFullyConnected 324 567 :=> SpecLeakyRelu (567,1,1) :=> SpecFullyConnected 567 486 :=>
% SpecLeakyRelu (486,1,1) :=> SpecFullyConnected 486 36 :=> SpecReshape (36,1,1) (18,2,1) :=>
% SpecNNil2D18x2
%

@@

\subsection{Deep Q-Learning}
\MS[t]{explain here, alg in appendix?}

@@

\begin{table}[t]
  \centering
  \footnotesize
  \begin{tabular}{lP{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}}
    \toprule
    Parameter                      & $\alpha$   & $\gamma_{0}$ & $\gamma_{1}$ & $\gamma$ & $\epsilon$ & $p_{exp}$ & $\plearn$ \\
    \midrule
    Start value                    & 0.01       & 0.8          & 0.99/1.0     & 0.01     & 0.25       & 1.0       & 0.5       \\
    Exp. Smth. Rate                & 0.25       &              &              & 0.25     &            & 0.25      &           \\
    Exp. Smth. Steps in \(10^{3}\) & 50         &              &              & 100      &            & 100       &           \\
    Minimum value                  & \(5^{-5}\) &              &              & 0.001    &            & 0.005     &           \\
    \bottomrule
    \toprule
    Parameter                      & \multicolumn{7}{c}{Value}                                                                                  \\
    \midrule
    Workers Min. Exploration       & \multicolumn{7}{c}{0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20}                                         \\
    Training Batch Size            & \multicolumn{7}{c}{4}                                                                                      \\
    Target Network Update          & \multicolumn{7}{c}{Exp. smooth. with rate \(\gamma\) every \(C=100\) steps}                                \\
    N-Step                         & \multicolumn{7}{c}{5}                                                                                      \\
    Training Mini-Batch Size       & \multicolumn{7}{c}{\(m = 4\)}                                                                              \\
    Adam ANN Backpropagation       & \multicolumn{7}{c}{\(\alpha = 0.005, \beta_{1}=0.9, \beta_{2}=0.999, \epsilon=10^{-8}, \lambda=0.001\)}      \\ % OptAdam 0.005 0.9 0.999
% 1e-8 1e-3,
        Replay Memory Size         & \multicolumn{7}{c}{\(10800\)}                                                                              \\
    \bottomrule
  \end{tabular}
  \caption{\label{tbl:params}Overview of Parameter Setup.}
\end{table}
%
Table~\ref{tbl:params} presents an overview of the parameter setup. The parameters \(\alpha\),
\(\gamma\) and \(p_{exp}\) are exponentially decayed as indicated.% \footnote{The decay is defined as
  % \(0.05 ^{(t/350k)}\), where \(t\) is the current period}.
\MS[t]{TODO} As further comparison we implement the deep Q-Learning algorithm proposed
by~\cite{mnih2015human}, where for comparison reasons we also integrate n-Step learning, the
experience replay memory adaptions and independent agents. \MS[t]{Discount factor in table?}

@@

@@

\begin{algorithm}[h!]
  \begin{algorithmic}[1]
    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly,
    set an exploration rate \(0 < \pexp \leqslant 1\), exponential smoothing learning rates
    \(0 < \alpha_{\max}, \alpha, \gamma < 1\) and discount factor \(0 < < \gamma_{1} < 1\)
    \While{the stopping criterion is not fulfilled} \ForEach{independent agent
      \(i \in \agent_{1},\agent_{2},\ldots,\agent_{n}\)} \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      With probability \(\frac{\pexp}{n}\) choose action \(a_{t,i}\) randomly or otherwise one of
      the set defined by
      \(\min_{a \in \mathcal{A}(i)}\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{})\).
    \end{minipage}
    \EndFor{}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
      Carry out action \(a_{t}=(a_{t,\agent_{1}},\ldots,a_{t,\agent_{n}})\), observe reward
      \(r_{t}\) and resulting state \(s_{t+1}\). Store experience
      \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay memory \(M\).
    \end{minipage}
    \If{\(t \mod{\nstep} \equiv 0\)}
    \State{Reset gradient: \(d\netWorker{} \gets 0\)}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      Sample \(m\) random mini-batches of \(\nstep\) consecutive experiences from \(M\) numbered by
      \(1 \leqslant i \leqslant \nstep\), each with
      \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i+1})\).
    \end{minipage}
    \ForEach{mini-batch}
    \ForEach{experience, starting with the latest seen experience \(\nstep\)}
    \If{experience \(\nstep\) or \(\neg \isRandAct{i}\) or  \(\pexp > \plearn\)}
      \State{\(\target_{\gamma1} \gets \min_{a} \X^\pol_{\gamma_1}(s_{i+1},a; \netTarget{})\)}
    \EndIf
    \State{\(\target_{\gamma_{1}}  \gets r_{i} + \gamma_{1} \target_{\gamma_{1}} \)}
    \State{\(d\netWorker{} \gets d\netWorker{} +
      \partial{(\target_{\gamma_{1}} - \X^\pol_{\gamma_{1}}(s_{i+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
    \EndFor{}
    \EndFor{}
    \State{Perform update of \(\netWorker{}\) using \(d\netWorker{} / (m \cdot \nstep)\)}
    \EndIf{}
    \State{Every \(C\) steps exponential smoothly set target network:
    \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)}
    \State{Set \(s \gets s'\), \(t \gets t+1\) and decay parameters}
    \EndWhile{}
  \end{algorithmic}
  \caption{\label{alg:ql}Deep Q-Learning}
\end{algorithm}

>>>>

\paragraph{Demand.} The incoming orders are placed in the order pool with a due date slack of
\dds{} periods, that is each order arriving within period $t$ will be due at the end of period
$t+\dds{}$. The due date slack of \(\dds{}\) was chosen to (i) ensure sufficient time between the
first occurrence of orders in the order pool and the latest possible release of orders and (ii) to
have a clear cause and effect relationship between the order release decision and the cost
performance in the analysis.
%
To simulate medium and high variability of the demand process the interarrival time between
consecutive incoming orders is either drawn from an exponential ($\Exp$) or an uniform ($\Unif$)
distribution, which are adjusted to yield the desired bottleneck utilisation level of either $70\%$
with $\Exp(135)$ and $\Unif(95,175)$, $80\%$ with $\Exp(118)$ and $\Unif(78,158)$, or $90\%$ with
$\Exp(105)$ and $\Unif(65,145)$. These values were chosen as they comprise the non-linear
relationship between flow times and high utilisation levels.

@@

\paragraph*{Neural Network Setup.}
%
Due to the vast possibilities the neural network architecture was manually optimised, starting with
less stochastic systems and aiming for a coffin shape that can be used in all setups. During this
process we constantly adapted and reevaluated the best network on various production system
configurations. The resulting architecture depends on the number of features \(\anninp\) and the
number of output nodes \(\annout\), which itself depends on the number of actions
\(\acts = \size{\States}\), the independent agent configuration, and the number of functions that
are approximated. For the later, in case of \ARA{} this is \(2\), as we represent both functions
\(\X^\pol_{\gamma_{0}}\) and \(\X^\pol_{\gamma_{1}}\) with the same neural network, while deep
Q-Learning \citep[cf.][]{mnih2015human} only approximates the function \(Q_{\gamma}^\pol\).
%
The resulting feedforward deep ANN architecture is given as %
\(\anninp %
\rightarrow 1.75 \anninp % \rightarrow \relu
\rightarrow 1.5 \anninp % \rightarrow \relu
% \rightarrow 2 \annout % \rightarrow \relu
\rightarrow \annout %
% \rightarrow   \leakyTanh%
\), where each arrow symbolises a fully-connected layer with a leaky rectified linear unit
(\(\relu\)) as activation function, which leaks values smaller than \(0\) with rate \(0.02\).
%
Note
that we explicitly do not use an output activation function. The idea is that we do not want to
restrict the approximations to values inside this range. This is especially important for
Q-Learning, as the range of the state values are difficult to forecast.
%
The actual values \(\anninp\) and \(\annout\) depend on the used algorithm. In our case using 6
independent agents we gain \(\anninp=324\) and \(\annout=36\).
%
% SpecFullyConnected 324 567 :=> SpecLeakyRelu (567,1,1) :=> SpecFullyConnected 567 486 :=>
% SpecLeakyRelu (486,1,1) :=> SpecFullyConnected 486 36 :=> SpecReshape (36,1,1) (18,2,1) :=>
% SpecNNil2D18x2
%

@@

\subsection{Deep Q-Learning Algorithm}

Furthermore, we adapt the deep Q-Learning algorithm of \citet{mnih2015human} with integrated
extensions as described above, i.e.\@ the adaptions to n-step learning, multiple workers,
independent agents, scaling of the output, smoothly adapting the target network. The resulting
Algorithm~\ref{alg:ql} is provided in the Appendix.

To ensure comparability we use the same neural network setup as for \ARA{}, except that we halve the
number of output nodes, as the Q-Learning algorithm only approximates one state-action value
function. Therefore the output nodes \(\annout\) are \(18\), while the input nodes \(\anninp=324\)
stay the same.

@@

\begin{table}[t]
  \centering
  \footnotesize
  \begin{tabular}{lP{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}}
    \toprule
    Parameter                      & $\alpha$   & $\gamma_{0}$ & $\gamma_{1}$ & $\gamma$ & $\epsilon$ & $p_{exp}$ & $\plearn$ \\
    \midrule
    Start value                    & 0.01       & 0.8          & 0.99/1.0     & 0.01     & 0.25       & 1.0       & 0.5       \\
    Exp. Smth. Rate                & 0.25       &              &              & 0.25     &            & 0.25      &           \\
    Exp. Smth. Steps in \(10^{3}\) & 50         &              &              & 100      &            & 100       &           \\
    Minimum value                  & \(5^{-5}\) &              &              & 0.001    &            & 0.005     &           \\
    \bottomrule
    \toprule
    Configuration                  & \multicolumn{7}{c}{Value}                                                                                  \\
    \midrule
    Workers Min. Exploration       & \multicolumn{7}{c}{\(0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20\)}                                         \\
    Target Network Update          & \multicolumn{7}{c}{Exp. smooth. with rate \(\gamma\) every \(C=100\) steps}                                \\
    N-Step                         & \multicolumn{7}{c}{\(5\)}                                                                                      \\
    Training Mini-Batch Size       & \multicolumn{7}{c}{\(4\)}                                                                              \\
    Adam ANN Backpropagation       & \multicolumn{7}{c}{\(\alpha = 0.005, \beta_{1}=0.9, \beta_{2}=0.999, \epsilon=10^{-8}, \lambda=0.001\)}      \\ % OptAdam 0.005 0.9 0.999
% 1e-8 1e-3,
        Replay Memory Size         & \multicolumn{7}{c}{\(10800\)}                                                                              \\
    \bottomrule
  \end{tabular}
  \caption{\label{tbl:params}Overview of Parameter Setup.}
\end{table}
%
Table~\ref{tbl:params} presents an overview of the parameter setup. The parameters \(\alpha\),
\(\gamma\) and \(p_{exp}\) are exponentially decayed as indicated. Due to the vast possibilities we
manually optimised the values. The methodology was to investigate the values on very small
production system with just one machine and product due to the computational complexity of bigger
sized problems. The resulting values were copied to the problem at hand and further manually
optimised in pre-experimental runs. The actual used discount factors \(\gamma_{1}\) differ for each
method and thus are further specified in the result tables.
% \footnote{The decay is defined as \(0.05 ^{(t/350k)}\), where \(t\) is the current period}.
In a similar approach we reached the configuration values, where a further increase of the values
was hindered by the available computational resources.

For the experimental runs we used personal computers with Intel CPUs i7-7700 (16GB RAM) and i9-7900X
(32GB RAM). For each method we performed \(500.000\) learning steps, which took about \(2\)-\(3\)
days. For the same experimental setup the inputs/random variables are the same for each period.
After learning each order release method was evaluated on \(20\) demand streams, each with \(1000\)
periods of warm-up and \(7000\) periods of evaluation phase. Welch’s procedure was applied to
approximate the length of the warm-up period (see~\cite{law:simulationc}).

\paragraph{Medium Variability.} The results for medium variability in the system are presented in
Table~\ref{tbl:unif}, where a uniform demand interarrival time is used.

@@

      % For 10 replications

@@

\begin{algorithm}[h!]
  \begin{algorithmic}[1]
    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly,
    set exploration rate \(0 < \pexp \leqslant 1\), target network adaption rate
    \(0 < \gamma \ll 1\) and discount factor \(0 < \gamma_{1} < 1\)
    \While{the stopping criterion is not fulfilled} \ForEach{independent agent
      \(i \in \agent_{1},\agent_{2},\ldots,\agent_{n}\)} \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      With probability \(\frac{\pexp}{n}\) choose action \(a_{t,i}\) randomly or otherwise one of
      the set defined by
      \(\min_{a \in \mathcal{A}(i)}\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{})\).
    \end{minipage}
    \EndFor{}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
      Carry out action \(a_{t}=(a_{t,\agent_{1}},\ldots,a_{t,\agent_{n}})\), observe reward
      \(r_{t}\) and resulting state \(s_{t+1}\). Store experience
      \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay memory \(M\).
    \end{minipage}
    \If{\(t \mod{\nstep} \equiv 0\)}
    \State{Reset gradient: \(d\netWorker{} \gets 0\)}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      Sample \(m\) random mini-batches of \(\nstep\) consecutive experiences from \(M\) numbered by
      \(1 \leqslant i \leqslant \nstep\), each with
      \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i+1})\).
    \end{minipage}
    \ForEach{mini-batch}
    \ForEach{experience, starting with the latest seen experience \(\nstep\)}
    \If{experience \(\nstep\) or \(\neg \isRandAct{i}\) or  \(\pexp > \plearn\)}
      \State{\(\target_{\gamma1} \gets \min_{a} \X^\pol_{\gamma_1}(s_{i+1},a; \netTarget{})\)}
    \EndIf
    \State{\(\target_{\gamma_{1}}  \gets r_{i} + \gamma_{1} \target_{\gamma_{1}} \)}
    \State{\(d\netWorker{} \gets d\netWorker{} +
      \partial{(\target_{\gamma_{1}} - \X^\pol_{\gamma_{1}}(s_{i+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
    \EndFor{}
    \EndFor{}
    \State{Perform update of \(\netWorker{}\) using \(d\netWorker{} / (m \cdot \nstep)\)}
    \EndIf{}
    \State{Every \(C\) steps exponential smoothly set target network:
    \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)}
    \State{Set \(s \gets s'\), \(t \gets t+1\) and decay parameters}
    \EndWhile{}
  \end{algorithmic}
  \caption{\label{alg:ql}Deep Q-Learning}
\end{algorithm}
